---
title: "265-lp-04"
author: "Will Park 265"
date: "2025-04-09"
output:
  pdf_document: default
  html_document: default
---
```{r}
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(scales)
library(rvest)
library(stringr)
library(tidyverse)
```


```{r}
price_data <- read.csv("BitcoinPriceData.csv")
trend_data <- read.csv("BTCsent.csv")
```

### Task: Visualize Bitcoin Search Interest

- **Objective:** Visualize interest based on Google search trends 
- **Task Type:** Data Visualization  

#### Implementation desc. 
Used Google Trends data (2015–2025) to display public search behavior. 
This reveals changes in public attention over time and supports predictive modeling.

IMPORT
Imported BTCsent.csv, which includes yearly search volume index data for the term “Bitcoin” from 2015–2025.

TIDY
Data was already structured with Year and bitcoin.price.searches. No transformation was necessary.

UNDERSTAND LOOP
Explore: Used search trends to understand fluctuations in public attention.
Model: Not modeled, only visualized as raw trends.
Refine: Added labels and formatted lines to highlight interest patterns.

COMMUNICATE
Line + point plot communicates changes in public sentiment over time.

```{r}
ggplot(trend_data, aes(x = Year, y = bitcoin.price.searches)) +
  geom_line(color = "darkred", size = 1.2) +
  geom_point(color = "darkred", size = 1.2) +
  labs(
    title = "Bitcoin Search Interest (Google Trends, 2015–2025)",
    x = "Year",
    y = "Average Search Interest (0–100 scale)"
  ) +
  theme_minimal()
```


### Task: Visualize Average Yearly Bitcoin Price

- **Objective:** Analyze historical Bitcoin prices to establish a pricing trend for modeling.
- **Task Type:** Data Visualization  

#### Implementation desc. 
Extracted yearly average prices (2018–2025) from historical data and plotted trends.
This helps identify significant changes in market behavior and prepares data for forecasting.

IMPORT
Used btc_historical.csv, containing daily Bitcoin prices with timestamps.

TIDY
Renamed and converted the date field.

Extracted Year, filtered for 2018–2025.

Aggregated daily data into yearly averages.

UNDERSTAND LOOP
Explore: Investigated long-term price movement.
Model: Computed yearly averages using mean(Open).
Refine: Used dollar formatting and range restrictions for clarity.

COMMUNICATE
Line chart clearly communicates historical growth and trend stability/volatility.

```{r}

btc_data <- read_csv("btc_historical.csv")
btc_data <- btc_data %>% rename(Date = `Open time`)
btc_data$Date <- as.Date(btc_data$Date)
btc_data$Year <- year(btc_data$Date)

btc_yearly_avg <- btc_data %>%
  filter(Year >= 2018, Year <= 2025) %>%
  group_by(Year) %>%
  summarise(Avg_Open = mean(Open, na.rm = TRUE))

ggplot(btc_yearly_avg, aes(x = Year, y = Avg_Open)) +
  geom_line(color = "orange", linewidth = 1.2) +
  geom_point(color = "orange") +
  scale_y_continuous(
    limits = c(0, 110000),
    breaks = seq(0, 110000, by = 10000),
    labels = dollar_format()
  ) +
  scale_x_continuous(breaks = btc_yearly_avg$Year) +
  labs(
    title = "Average Yearly Bitcoin Price",
    x = "Year",
    y = "Average Opening Price (USD)"
  ) +
  theme_minimal()
```


### Task: Predicting Bitcoin Prices Based on Google Trends

- **Objective:** Forecast Bitcoin prices using public interest (search trends) and time (year).
- **Task Type:** Predictive Analytics  

#### Implementation desc. 
Applied linear and multivariate regression using search interest and year.Predicted future prices and visualized with confidence intervals.

IMPORT
Used both btc_yearly_avg and trend_data created above.

TIDY
Joined trend and price data by Year.

Selected necessary columns for modeling.

Created forecast years manually for multivariate model.

UNDERSTAND LOOP
Explore: Evaluated the linearity of the relationship between interest and price.
Model: Ran linear and multivariate regressions.
Refine: Used confidence intervals, layered plots, and future projections.

COMMUNICATE
Displayed both observed and forecasted prices.
Plotted regression fit and shaded confidence intervals.


```{r}
predict_df <- btc_yearly_avg %>%
  inner_join(trend_data, by = "Year") %>%
  select(Year, Avg_Open, bitcoin.price.searches)

# Linear regression: Avg_Open ~ bitcoin.price.searches
model <- lm(Avg_Open ~ bitcoin.price.searches, data = predict_df)
summary(model)

predict_df <- predict_df %>%
  mutate(Predicted_Price = predict(model, newdata = predict_df))

ggplot(predict_df, aes(x = bitcoin.price.searches)) +
  geom_point(aes(y = Avg_Open), color = "blue", size = 3) +
  geom_line(aes(y = Predicted_Price), color = "red", linewidth = 1.2) +
  labs(
    title = "Predicted vs Actual Bitcoin Prices Based on Google Trends",
    x = "Google Search Interest (0–100)",
    y = "Average Bitcoin Price (USD)"
  ) +
  theme_minimal()


# Multivariate regression with forecast
multi_model <- lm(Avg_Open ~ Year + bitcoin.price.searches, data = predict_df)

future_years <- data.frame(
  Year = 2026:2028,
  bitcoin.price.searches = c(30, 35, 40)
)

full_years <- bind_rows(
  predict_df %>% select(Year, bitcoin.price.searches),
  future_years
)

predictions <- predict(multi_model, newdata = full_years, interval = "confidence")
full_years <- bind_cols(full_years, as.data.frame(predictions))

full_years <- full_years %>%
  mutate(Data_Type = ifelse(Year <= max(predict_df$Year), "Actual", "Forecast"))

ggplot(full_years, aes(x = Year, y = fit, color = Data_Type)) +
  geom_line(size = 1.2) +
  geom_point(data = predict_df, aes(x = Year, y = Avg_Open), color = "blue", size = 3) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "gray80", alpha = 0.3) +
  labs(
    title = "Bitcoin Price Forecast Using Year + Google Trends",
    x = "Year",
    y = "Average Bitcoin Price (USD)"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("Actual" = "blue", "Forecast" = "red"))
```

### Task: Querying Crypto Fraud Cases

- **Objective:** Assess the rise in fraud cases  
- **Task Type:** Database Query  

**Implementation Description:**  
I attempted to query CryptoScamDB for scam reports but received a 502 server error. As a fallback, I used the BitcoinHeist dataset to summarize Bitcoin fraud cases by year. This query revealed peak activity around 2016–2018. The objective remained unchanged; only the data source was adjusted.

IMPORT
Used BitcoinHeistData.csv as a fallback after API failure.

TIDY
Filtered only fraudulent cases (label != "white").

Grouped by year.

UNDERSTAND LOOP
Explore: Identified periods of heightened fraud activity.
Model: Counted total cases per year.
Refine: Used basic column plot.

COMMUNICATE
Time series of fraud counts over time communicates urgency and risk pattern.

```{r}
fraud_data <- read_csv("BitcoinHeistData.csv")

# Summarize only fraudulent cases by year
fraud_summary <- fraud_data %>%
  filter(label != "white") %>%   # filter only fraud cases
  group_by(year) %>%
  summarise(Fraud_Cases = n())

# Plot
ggplot(fraud_summary, aes(x = year, y = Fraud_Cases)) +
  geom_col(fill = "darkred") +
  labs(
    title = "Reported Bitcoin Fraud Cases by Year",
    x = "Year", y = "Number of Fraud Cases"
  ) +
  theme_minimal()

```


### Task: Most Common Bitcoin Fraud Types

- **Objective:** Assess the rise in fraud cases  
- **Task Type:** Database Query  

#### Implementation Description  
To identify which fraud types are most common in the Bitcoin ecosystem, I grouped all fraudulent transactions (excluding those labeled "white") by fraud label and summarized their frequency.  
The plot clearly shows phishing and ransomware as the most reported fraud types. This supports the goal of understanding fraud prevalence across categories.

TIDY
Grouped by label, then sorted by frequency.

UNDERSTAND LOOP
Explore: Revealed the most dominant types of fraud.
Model: Count of each fraud type.
Refine: Ordered bars and rotated text for clarity.

COMMUNICATE
Bar chart shows which scams dominate the ecosystem.

```{r}
fraud_types <- fraud_data %>%
  filter(label != "white") %>%
  group_by(label) %>%
  summarise(Cases = n()) %>%
  arrange(desc(Cases))

ggplot(fraud_types, aes(x = reorder(label, -Cases), y = Cases)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Most Common Bitcoin Fraud Types",
    x = "Fraud Type", y = "Number of Cases"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Task: Predicting Bitcoin Fraud Based on Network Features and Market Growth

- **Objective:** Develop predictive models  
- **Task Type:** Predictive Modeling  

#### Implementation Description  
To explore whether fraud can be predicted based on address characteristics and market trends, I developed a logistic regression model using the BitcoinHeist dataset. I created a binary outcome variable (`is_fraud`) based on whether the `label` was “white” (legit) or a fraud type. I added annual Bitcoin growth as an external variable and modeled fraud probability using `length`, `weight`, and `growth`

TIDY
Created binary is_fraud column.

Merged external market data with fraud records.

UNDERSTAND LOOP
Explore: Hypothesized that address features + market growth affect fraud.
Model: Logistic regression with glm().
Refine: Output summary to evaluate variable significance.

COMMUNICATE
Model summary indicates predictors’ effects on fraud probability.

```{r}
# Load data
fraud_data <- read_csv("BitcoinHeistData.csv")

# Label fraud
fraud_data <- fraud_data %>%
  mutate(is_fraud = ifelse(label == "white", 0, 1))

# External feature: market growth
btc_growth <- tibble(
  year = c(2013, 2014, 2015, 2016, 2017),
  btc_price = c(800, 500, 600, 970, 4200),
  growth = c(NA, -37.5, 20.0, 61.7, 333.0)
)

# Merge into fraud dataset
fraud_data <- fraud_data %>%
  left_join(btc_growth, by = "year")

# Model data
model_data <- fraud_data %>%
  select(length, weight, growth, is_fraud) %>%
  filter(!is.na(growth))

# Fit logistic regression
model <- glm(is_fraud ~ length + weight + growth, data = model_data, family = binomial)
summary(model)

```

Additional notes and update to proposal:

Data set fault:
- The dataset btc_historical was lost and had to be replaced but the updated csv varied significantly from the previous version. The current code does NOT reflect this change and needs to be updated. 

Project overview:
- Explore the relationship between public interest in Bitcoin, its market value trends, and the prevalence of Bitcoin-related fraud. The project utilizes datasets from Google Trends, historical Bitcoin pricing, and the BitcoinHeist fraud database to:

Primary objectives:
- Data exploration: visualize interest over time, historical average prices, frequency and types of fraud cases
- Predictive modeling: forecast prices based on search interest and time and predict fraud probability using transaction features and market context

